---
title: "Meteorite Landings"
author: "Jasmine Dumas (jasmine.dumas@gmail.com)"
date: "2017-03-11"
output: md_document
---

# Meteorite Landings

## Summary

This comprehensive data set from The Meteorological Society contains information on all of the known meteorite landings. 

## Analysis

The goal of this analysis is to create an insightful and interactive data visualization which draws some inference about meteorite landing locations, mass, and year of occurrence. 
```{r, message=FALSE, warning=FALSE}
library(ggplot2)
library(RSocrata)
library(devtools)
library(gganimate)
library(magick)
library(dplyr)
library(ggmap)
library(gapminder)
library(animation)
library(lubridate)
```

```{r}
data_in <- read.socrata("https://data.nasa.gov/Space-Science/Meteorite-Landings/gh4g-9sfh")
```

```{r}
head(data_in)
```

### Clean up column names
```{r}
names(data_in)[c(3, 4, 5, 6, 8, 9)] <- c("name_type", "class", "mass", "found", "latitude", "longitude")
names(data_in)
```
### Fix the year by removing MM-DD
```{r}
data_in$year_fix <- year(data_in$year)
```
### Explore the variables by class distribution
```{r}
data_in %>% group_by(class) %>% 
            summarise(n = n()) %>% 
            arrange(desc(n))
```

```{r}
data_in %>% group_by(name_type) %>% 
            summarise(n = n()) %>% 
            arrange(desc(n))
```

### Filter out erroneous latitudes and longitudes
```{r}
data_in_clean <- data_in %>% filter(latitude >= -90 & 
                                    latitude  <= 90 & 
                                    longitude >= -180 &
                                    longitude <= 180  & 
                                    !is.na(mass), 
                                    !is.na(year))
```

### Clean up method of whether the meteorite was Found or not
```{r}
data_in_clean$found <- if_else(data_in_clean$found == "Fell", "No", "Yes")
```


## Visualizations

### Set up some options for plotting
```{r}
theme_set(theme_minimal())

options(scipen=10000)
```

### Static display
```{r}
p <- ggplot() +
    borders("world", colour="gray50") +
    geom_point(aes(x = longitude, y = latitude, color = found, frame = year_fix), data = data_in_clean, alpha = .5) +
    labs(title = "Worldwide Meteorite Landings ", 
       x = "", y = "") +
    theme(legend.position="bottom", 
        axis.text.x = element_blank(),
        axis.text.y = element_blank(),
        axis.ticks = element_blank())
p
```


```{r, fig.show = "animate"}
gganimate(p, "meteorite-landings.gif")
```

## Insights

* Dominant areas where meteorites were found are: Australia, USA, and Antarctica


## Resources

*I reused some code and modified examples from:*

1. [https://rpubs.com/MVellinger/Gibeon](https://rpubs.com/MVellinger/Gibeon)

2. [http://relevantmisc.com/ggplot/instagram/2016/02/14/ig-over-time/](http://relevantmisc.com/ggplot/instagram/2016/02/14/ig-over-time/)

3. [https://rpubs.com/sjackman/gapminder-gganimate](https://rpubs.com/sjackman/gapminder-gganimate)

*Thanks to [Bob Rudis](https://twitter.com/hrbrmstr/status/840622466707922944) for helping me to get ImageMagick installed (Twitter is a powerful resource)*